%
% File acl2014.tex
%
% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
%%
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2014}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\DeclareMathOperator*{\argmax}{arg\,max}
%\DeclareMathOperator*{\argmin}{arg\,min}

\usepackage{graphicx}



%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.
%\input{defs}

\title{Parameter Agreement Training with and without Parallel Data}

%\author{First Author \\
%  Affiliation / Address line 1 \\
%  Affiliation / Address line 2 \\
%  Affiliation / Address line 3 \\
%  {\tt email@domain} \\\And
%  Second Author \\
%  Affiliation / Address line 1 \\
%  Affiliation / Address line 2 \\
%  Affiliation / Address line 3 \\
%  {\tt email@domain} \\}
%\date{}

\input{math_definition.tex}

\begin{document}
\maketitle
\begin{abstract}
%% this doesn't make a lot of sense
%We present a simple approach that encourages jointly trained conditional models to agree on their parameter values. In contrast to methods that encourage agreement on inferences, our method can be applied in a wider range of settings.
%We demonstrate that our method, called parameter agreement training, leads to state-of-the-art accuracies on word alignment, and significantly improves accuracy on back-transliteration with no parallel data -- almost halving the difference between the previous best method and using parallel data.
Alignment is a crucial subtask for many larger NLP problems: Word alignments are the backbone for extracting translation rules used in statistical machine translation; phoneme alignments are used in transliteration from Japanese katakana to English and vice-versa. The popular models used in these tasks are both \emph{generative} and \emph{directed}, that is, they define a directed sequence of events by which one side of the sequence pair produces the other side. For example, the IBM models for word alignment \newcite{ brown+alii:1993}, describe a stochastic process by which the target sentence is generated from the source sentence via word alignments. For word alignment, the generative processes allow a target word to align to \emph{at most} one source word which causes the generative processes to produce different alignments in the source-target and target-source directions. For both word alignment and phoneme alignment, the models in each direction are trained independently of each other. While this is not so much a problem for phoneme alignment, word alignment models end up producing different alignments in each direction and to correct for the errors that arise in each direction, we typically resort to ad-hoc alignment symmetrization approaches \cite{koehn+:2003}. A second problem for training alignment models is the need for parallel data. For example, in \newcite{KG98}, the authors use \marginpar{insert number} of parallel English-Katana sequences to train the alignment models. While large amounts of parallel data needed for word alignment is readily available for many language pairs, acquiring parallel phoneme sequences for transliteration can be very challenging. There has been previous work on solving the problem of model disagreement in word alignment~\cite{liang+:2006:align} and training phoneme alignment models from large amounts of monolingual data~\cite{RK09}. However, in \newcite{liang+:2006:align}, the authors do not optimize a clear objective function and their approach cannot be applied to training alignment models without parallel data. ~\newcite{RK09} improve monolingual phoneme alignment over the baseline but their approach does not encourage model agreement, which can improve alignment accuracy. In our paper, we present a single approach that achieves both model agreement during training, and can be used in non-parallel settings. Our approach, \emph{Parameter Agreement Training}, encourages model agreement by adding a regularization term that penalizes disagreement between the (magnitudes of the ?) model parameters in each direction. Our regularizer is convex, allowing us to efficiently optimize the objective function with the EM algorithm. We apply our techniques to word alignment and phoneme alignment. Using our approach, we are able to significantly improve over ~\newcite{RK09}. In word alignment, our results compare favorably with results from previous approaches for model agreement.  

\end{abstract}

\section{Introduction}
\label{sec:introduction}
\input{introduction.tex}

\section{Background}
\label{sec:background}
\input{background.tex}

\section{Parameter Agreement Training}
\label{sec:method}
\input{method.tex}


\section{Transliteration}
\label{sec:transliteration}
\input{transliteration.tex}

\section{Word Alignment}
\label{sec:alignment}
\input{alignment.tex}

\section{Conclusion}
We have presented Parameter Agreement Training (PAT) - an approach for jointly training two conditional models that encourages agreement on parameter values.
We derived an optimization procedure based on the EM framework, and argued that it is efficiently solvable due to the concavity of our agreement regularizer.
In contrast to previous work on agreement training, we have demonstrated that our approach can be successfully applied even without parallel data.


\bibliographystyle{acl2014}
\bibliography{thesis}

\end{document}
